__Zero-Inflated Models (ZIM) for Count Time Series with Excess Zeros__

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/ZIM)](https://CRAN.R-project.org/package=ZIM)
[![CRAN RStudio mirror overall downloads](http://cranlogs.r-pkg.org/badges/grand-total/ZIM)](http://www.r-pkg.org/pkg/ZIM)
![CRAN Downloads month](http://cranlogs.r-pkg.org/badges/ZIM)

## Overview

Analyze count time series with excess zeros. Two types of statistical models are supported: Markov regression by Yang et al. (2013) and state-space models by Yang et al. (2015). They are also known as observation-driven and parameter-driven models respectively in the time series literature. The functions used for Markov regression or observation-driven models can also be used to fit ordinary regression models with independent data under the zero-inflated Poisson (ZIP) or zero-inflated negative binomial (ZINB) assumption. Besides, the package contains some miscellaneous functions to compute density, distribution, quantile, and generate random numbers from ZIP and ZINB distributions.

## Installation

### Stable Version

Install stable version from CRAN:

```{r, eval=FALSE}
install.packages("ZIM")
```

### Development Version

Install development version from GitHub:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("biostatstudio/ZIM")
```

## References

- Ming Yang, Gideon K.D. Zamba, Joseph E. Cavanaugh (2013). Markov Regression Models for Count Time Series with Excess Zeros: A Partial Likelihood Approach. _Statistical Methodology_, 14, 26-38.

- Ming Yang, Joseph E. Cavanaugh, Gideon K.D. Zamba (2015). State-Space Models for Count Time Series with Excess Zeros. _Statistical Modelling_, 15, 70-90.

